Improving the Image Acquisition Rate of an Atomic Force Microscope Through Sub-sampling and Reconstruction

### 2019-06-19 latex/afm-cs-main.tex
Final touches before sean and lucy see it.

\begin{document}
\title{Improving the Image Acquisition Rate of an Atomic Force
  Microscope Through Compressive Sampling}

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}

@@

In AFM, implementing a truly random pattern requires that the tip be
engaged with the surface to collect a measurement, lifted and moved to
the new location, and engaged again. Because the re-engagement process
is typically slow, this can lead to excessively long image acquisition
times and negate the gains from undersampling \cite{andersson2012non}.
One discrete undersampling scheme, called a $\mu$-path pattern, was
proposed in \cite{maxwell2014compressed} to improve sampling
efficiency. The $\mu$-path pattern consists of randomly placed short,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern) and
is designed to balance the randomness needed to ensure good
reconstruction with continuous scanning to reduce the number of tip
engagements.

\textbf{Contributions}
Our previous work on the $\mu$-path pattern using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve significant scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. This
paper builds upon that earlier work in several significant ways.

Much of the prior work on undersampled AFM
images is based purely on simulations \cite{han_optimal_2018,
  Luo_nano_2015, oxvig_structure_2017, jensen_reconstruction_2013,
  maxwell_acc_2014, Chen_2012jx} and consequently, there is little discussion of the physical practicality of the proposed methods. For example, most estimates of the potential time savings of sub-sampling techniques are not informed by experiments. Often, a discussion of the harmonic content of the proposed reference signals is typically neglected \cite{luo_continuous_2019, Chen_2012jx,han_optimal_2018}, even though this is a critical concern in AFM. As we show in Section~\ref{sec:control}, the discontinuous nature of the $\mu$-path pattern essentially requires that the control system be to filter the higher frequency content of the step inputs between $\mu$-paths.

Despite this, we show in Section~\ref{sec:results} that $\mu$-path scanning provides a non-trivial speed increase compared to a full quality raster scan. 
% % ------------------ TKTKTKT ------------------
% \todo{Strike this?} First, we demonstrate performance using both
% randomly-placed $\mu-$paths and designed $\mu-$paths that are
% optimized to minimize the expected reconstruction error given some
% initial estimate of the frequency structure of the sample. These
% schemes are described in Sec. \ref{sec:samplingScheme}.
% % --------------------------
We also introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec.
\ref{sec:reconstructionMethods}. Section \ref{sec:control} describes
improvements to the control algorithms compared to our preliminary
implementation in ~\cite{braker_hardware_2018}. We then demonstrate
the entire scheme from sampling through reconstruction using
experiments on a commercial AFM (Agilent 5500). The experimental setup
and implementation details are described in Secs.
\ref{sec:experimentalSetup} and \ref{sec:implementation}\footnote{The
  software portion of the implementation and image reconstruction
  described in this paper can be found at
  \url{https://github.com/rabraker/AFM_CS}}. Experimental results are
provided in Sec. \ref{sec:results}. Finally, we provide a few
concluding remarks in Sec. \ref{sec:conclusions}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures-SBA/random_mask.pdf}
  \caption{An example of the horizontal $\mu$-path sampling pattern
    for a 256$\times$256 pixel image with 35 pixels in each short scan
    and a total of 20\% pixels scanned.}
  \label{fig:mu_mask}
\end{figure}
% =========================================================================
\subsection{Reconstruction methods}
\label{sec:reconstructionMethods}
Compressive sensing (CS) is a signal processing technique which aims
at signal reconstruction from a relatively small (sub-Nyquist limit)
number of measurements \cite{carmi2014compressive}. It takes advantage
of the approximate sparsity of real-world signals, that is, that many
coefficients of such signals are close to zero when represented in an
appropriate basis. CS methods seek the true image signal
$x\in\mathbb{R}^n$ from the observation equation,
\begin{equation}\label{op:observation}
  y = R x = RU\eta,
\end{equation}
\noindent where $y\in\mathbb{R}^m$ is the observation vector, $R$ is
an $n\times m$ matrix defining the measurements, $U$ is an $m\times m$
sparsity basis and $\eta$ is the sparse representation of $x$ in the
domain of $U$. In general, $n\ll m$. In an imaging application, where
the underlying signal is a matrix $X\in\mathbb{R}^{h\times s}$, we
take $x=\text{vec}(X)$, where the $\text{vec}(\cdot)$ operator stacks
the columns of a matrix.

In the AFM application, the probe can only measure a single pixel at a
time. Thus the rows of $R$ are a subset of the rows of an $m$ by $m$
identity matrix. Ideally, the sparsity basis and the measurement
matrix $R$ will have a low \textit{mutual coherence}, a measure that
describes how each of the rows of $R$ (the measurements) ``spreads
out'' in the domain of $U$ \cite{candes2007sparsity}. In the
following, we assume that $U^TU=I$, though in the simulation and
experimental results, we take $U^T$ as the Discrete Cosine Transform
(DCT). The DCT basis generally keeps a good balance between achieving
a low mutual coherence between $U$ and the $\mu$-path sensing matrices
$R$ and providing a high sparsity for typical AFM sample images.

@@

Although BPDN is effective in the general setting, using it for
reconstruction from horizontal $\mu$-path samples often yields
artifacts in the vertical direction (that is, the direction orthogonal
to the $\mu$-path scans) leading to strong discontinuities in the
image. These artifacts grow more prominent as the length of the
$\mu$-paths is increased \cite{maxwell2014compressed}, because these
longer paths lead to increased mutual coherence between the
measurement matrix $R$ and the sparsity basis $U$. However, longer
$\mu$-paths means fewer $\mu$-paths and thus shorter scan times.
\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
                                   & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
                                                          & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
                                         & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
                                                                          & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
                                                                          % ------------------------------
                                                                          % Row
                                                                          % 2
                                                                          % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
                                   & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
                                                          & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
                                         & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
                                                                          & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, 40 pixel long $\mu$-paths with 25\%
    sampling. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_v x=\text{vec}(\nabla_v X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Section 6.1,]{becker_nesta_2011}).
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
    %     (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}.

@@

Implementing the $\mu$-path scheme involves operating the AFM in
several distinct stages. In the $XY$-direction, the system transitions
from tracking a step command in the transition to a new
measurement location to tracking a short ramp during the actual $\mu$-path scan.
For our simple integral controller, this only involves changing the reference signals
$X_{ref}$ and $Y_{ref}$. In the $Z$-direction, the system must
transition between tip descent, surface scanning, and tip retraction. F

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. In the following, $r_X$ and $r_Y$ refer to
the references for the $X$ and $Y$ axes while $r_{Z,s}$ and $\rzup$
refer to the $Z$-axis setpoints during scanning and retraction,
respectively.

\begin{table}
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $r_X$,$r_Y$ & $r_Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}
Two cycles of this sequential process are illustrated by the time
series in Fig.~\ref{fig:cs_cycle}. The following
subsections describe in more detail several features of the CS
scanning process.

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1\textwidth]{figures/AFM_loop_xyz.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup
    is active for decision index $i\in\{0,1,5\}$, which correspond to
    states such that $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
\subsection{The pre-scan}
After the tip-descent, we begin scanning in a ``pre-scan'' phase
(state (3)), which is the turquoise portion of the trajectories in Fig.~\ref{fig:cs_cycle}.
There are two reasons for this. First, we have observed that there is less noise in the deflection signal while the $XY$-stage is in motion compared to sitting still. More importantly, it lets us bring the $X$-axis to its scanning velocity at the same time the transients from the $Z$-axis die out. Without this, we must extend the $\mu$-path by a distance equal to the steady-state error following a ramp. For an arbitrary scan speed, this means the $\mu$-path scan would need an extra 
\begin{equation}
  N = \left\lceil\lim_{z\rightarrow 1} \frac{1-H_x}{z-1}\right\rceil
\end{equation}
time steps, where $H_x$ is closed-loop transfer function for the $x$-axis, including $F_X$.
For our system $N=111$. By initiating the scan while the $Z$-axis settles, we eliminate much of
this overhead.

@@

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
The SSIM and PSNR have become reasonably popular in studies comparing
simulated AFM image reconstructions \cite{oxvig_structure_2017,
  han_optimal_2018, Luo_nano_2015}. However, it remains somewhat of an
open question how to best compare experimental AFM images. While we
use the SSIM and PSNR in Section~\ref{sec:results}, we believe those
numbers should be interpreted with some caution. In our experimental
setup, actuation in the $XY$-plane happens via the nPoint piezo stage
and the $XY$-direction of the original piezo scanner is uncontrolled.
This leads to substantial drift between images. To illustrate this, we
took a sequence of 6 raster scans of a sample grating, each at a 1~Hz
scan rate over a 5 micron by 5 micron area. The left image of
Fig.~\ref{fig:baseline_errors} shows the error between the first and
 sixth image. The PSNR and SSIM between
images the first and second images are \baseline{one2twoNoAlgnPsnr} and
\baseline{one2twoNoAlgnSsim}; between the first and sixth are
\baseline{one2sixNoAlgnPsnr} and \baseline{one2sixNoAlgnSsim},
respectively, even though on their own each image has good quality.

To mitigate this affect, we use an image alignment technique based on
the 2D cross-correlation between each image. Essentially, this takes a
sub-slice of image 1 (here, a 25 pixel inset on each edge), and finds
the offset in the other image that produces the smallest error. The
result of this procedure are shown in the right image of
Fig.~\ref{fig:baseline_errors}.
Although the situation is improved compared to the un-aligned images,
there is still noticeable stretching (the upper and lower portions appear considerably worse than the middle). Notably however, with the alignment technique, the SSIM is
\baseline{one2twoAlgnSsim} for both sets of images and
while the PSNR is increases to \baseline{one2twoAlgnPsnr} and \baseline{one2sixAlgnPsnr}
respectively.

@@

\section{Experimental Results}\label{sec:results}
We compared $\mu$-path scanning to two scenarios. First, we compare it
to full resolution raster scans at several scan rates in Section~\ref{sec:cp_full}.
Second we compare to course raster scans in Section \ref{sec:cp_subline}.

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/cs_raster_pixel_rows_6-5-2019.svg}
  \caption{Rows of pixels, as indicated by the red lines in
    Fig.~\ref{fig:resultsF1_images}. For clarity, not all images are
    included.}
  \label{fig:pixel_rows}
\end{figure}
For each raster scan and each CS scan, we computed the SSIM and PSNR
 using the 1~Hz scan as the master and
also computed the RDI metric. These are plotted against total acquisition time
in Fig.~\ref{fig:time_ssim} and ~\ref{fig:time_psnr}.
Unfortunately, the SSIM and PSNR metrics are not particularly
illuminating. There are several inconsistencies. For example, it makes no sense
the 1 Hz CS scan at 25\% has a \emph{worse} PSNR and SSIM than 5 Hz version;
nor does it make sense that the 8~Hz full raster scan has a worse SSIM and PSNR than
the 10~Hz full raster scan.

@@

% \begin{figure}
%   \includesvg[width=1\columnwidth]{figures/subline_vs_cs.svg}
%   \caption{(left) A 128 line raster scan interpolated onto 512
%   lines. (right) The CS scan with the nearest time.}
%   \label{fig:subline}
% \end{figure}

\section{Conclusions}\label{sec:conclusions}
The main conclusion of this chapter is that with CS, it seems that we
can acquire images faster yet (putatively) with less specimen damage
than is possible while raster scanning a full image. On the other hand, it appears that reconstructing a course raster scan makes more sense for faster scan rates. This is notable, since a course raster scan only requires an additional step in post-processing. It remains to be seen if this conclusion holds for a wide range of instruments.

>>>>

\begin{document}
\title{Improving the Image Acquisition Rate of an Atomic Force
  Microscope Through Sub-sampling and Reconstruction}

@@

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}.

@@

In AFM, implementing a truly random pattern requires that the tip be
engaged with the surface to collect a measurement, lifted and moved to
the new location, and engaged again.
Because the re-engagement process
is typically slow, this can lead to excessively long image acquisition
times and negate the gains from undersampling \cite{andersson2012non}.
In \cite{maxwell2014compressed}, we introduced a discrete undersampling scheme,
called a $\mu$-path pattern, which consists of short randomly placed,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern). It
is designed to balance the randomness needed (to ensure good
reconstruction) with continuous scanning (to reduce the number of tip
engagements).

Our previous work on the $\mu$-path pattern using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are:
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype, \cite{braker_hardware_2018} we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an affect which neither our prior experimental work nor the numerous simulation studies have accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to course raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a course raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{Luo_nano_2015,maxwell_acc_2014} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=0.65\columnwidth]{figures-SBA/random_mask.pdf}
  \caption{An example of the horizontal $\mu$-path sampling pattern
    for a 256$\times$256 pixel image with 35 pixels in each short scan
    and a total of 20\% pixels scanned.}
  \label{fig:mu_mask}
\end{figure}
% =========================================================================
\section{Reconstruction methods}
\label{sec:reconstructionMethods}
Compressive sensing (CS) is a signal processing technique which aims
at signal reconstruction from a relatively small (sub-Nyquist limit)
number of measurements \cite{carmi2014compressive}. It takes advantage
of the approximate sparsity of real-world signals, that is, that many
coefficients of such signals are close to zero when represented in an
appropriate basis. CS methods seek the true image signal
$x\in\mathbb{R}^n$ from the observation equation,
\begin{equation}\label{op:observation}
  y = R x = RU\eta,
\end{equation}
\noindent where $y\in\mathbb{R}^m$ is the observation vector, $R$ is
an $m\times n$ matrix defining the measurements, $U$ is an $n\times n$
sparsity basis and $\eta$ is the sparse representation of $x$ in the
domain of $U$. In general, $m\ll n$. In an imaging application, where
the underlying signal is a matrix $X\in\mathbb{R}^{h\times s}$, we
take $x=\text{vec}(X)$, where the $\text{vec}(\cdot)$ operator stacks
the columns of a matrix.

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, 40 pixel long $\mu$-paths with 25\%
    sampling. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

In the AFM application, the probe can only measure a single pixel at a
time. Thus the rows of $R$ are a subset of the rows of an $n$ by $n$
identity matrix. Ideally, the sparsity basis and the measurement
matrix $R$ will have a low \textit{mutual coherence}, a measure of
how each of the rows of $R$ (the measurements) ``spreads
out'' in the domain of $U$ \cite{candes2007sparsity}. In the
following, we assume that $U^TU=I$, though in the simulation and
experimental results, we take $U^T$ as the Discrete Cosine Transform
(DCT). The DCT basis generally keeps a good balance between achieving
a low mutual coherence between $U$ and the $\mu$-path sensing matrices
$R$ and providing a high sparsity for typical AFM sample images.

@@

Although BPDN is effective in the general setting, using it for
reconstruction from horizontal $\mu$-path samples often yields
artifacts in the vertical direction (that is, the direction orthogonal
to the $\mu$-path scans) leading to strong discontinuities in the
image. These artifacts grow more prominent as the length of the
$\mu$-paths is increased \cite{maxwell2014compressed}, because these
longer paths lead to increased mutual coherence between the
measurement matrix $R$ and the sparsity basis $U$. However, longer
$\mu$-paths means fewer $\mu$-paths and thus shorter scan times.

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_v x=\text{vec}(\nabla_v X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Sec. 6.1,]{becker_nesta_2011}).
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
    %     (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}.

@@

Implementing the $\mu$-path scheme involves operating the AFM in
several distinct stages. In the $XY$-direction, the system transitions
from tracking a step command in the transition to a new
measurement location to tracking a short ramp during the actual $\mu$-path scan.
In the $Z$-direction, the system must
transition between tip descent, surface scanning, and tip retraction.
Two cycles of this sequential process are illustrated by the time
series in Fig.~\ref{fig:cs_cycle}. The following
subsections describe in more detail several features of the CS
scanning process.

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task.
In the following, $r_X$ and $r_Y$ and $r_Z$ refer to
the references for the $X$ and $Y$ axes while $r_{Z,s}$ and $\rzup$
refer to the $Z$-axis setpoints during scanning and retraction,
respectively.

\begin{table}[h!]
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $r_X$,$r_Y$ & $r_Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1\textwidth]{figures/AFM_loop_xyz.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup
    is active for decision index $i\in\{0,1,5\}$, which correspond to
    states such that $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
\subsection{The pre-scan}
After the tip-descent, we begin scanning in a ``pre-scan'' phase
(state (3)), which is the turquoise portion of the trajectories in Fig.~\ref{fig:cs_cycle}.
There are two reasons for this. First, we have observed that there is less noise in the deflection signal while the $XY$-stage is in motion compared to sitting still. More importantly, it lets us bring the $X$-axis to its scanning velocity at the same time the transients from the $Z$-axis die out. Without this, we must extend the $\mu$-path by a distance equal to the steady-state error following a ramp. For an arbitrary scan speed, this means the $\mu$-path scan would need an extra 
\begin{equation}
  N = \left\lceil\lim_{z\rightarrow 1} \frac{1-H_x}{z-1}\right\rceil
\end{equation}
time steps, where $H_x$ is closed-loop transfer function for the $x$-axis.
For our system $N=111$. By initiating the scan while the $Z$-axis settles, we eliminate much of
this overhead.

@@

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
The SSIM and PSNR have become reasonably popular in studies comparing
simulated AFM image reconstructions \cite{oxvig_structure_2017,
  han_optimal_2018, Luo_nano_2015}. However, it remains somewhat of an
open question how to best compare experimental AFM images. While we
use the SSIM and PSNR in Sec.~\ref{sec:results}, we believe those
numbers should be interpreted with some caution. In our experimental
setup, actuation in the $XY$-plane happens via the nPoint piezo stage
and the $XY$-direction of the original piezo scanner is uncontrolled.
This leads to substantial drift between images. To illustrate this, we
took a sequence of 6 raster scans of a sample grating, each at a 1~Hz
scan rate over a 5 micron by 5 micron area. The left image of
Fig.~\ref{fig:baseline_errors} shows the error between the first and
 sixth image. The PSNR and SSIM between
the first and second images are \baseline{one2twoNoAlgnPsnr} and
\baseline{one2twoNoAlgnSsim}; between the first and sixth are
\baseline{one2sixNoAlgnPsnr} and \baseline{one2sixNoAlgnSsim},
respectively, even though on their own each image appears to have good quality.

To mitigate this affect, we use an image alignment technique based on
the 2D cross-correlation between each image. Essentially, this takes a
sub-slice of the master (first) image (here, a 25 pixel inset on each edge), and finds
the offset in the other image that produces the smallest error. The
result of this procedure are shown in the right image of
Fig.~\ref{fig:baseline_errors}.
Although the situation is improved compared to the un-aligned images,
there is still noticeable stretching (the upper and lower portions appear considerably worse than the middle). Notably however, with the alignment technique, the SSIM is
\baseline{one2twoAlgnSsim} for both the second and sixth image images and
the PSNR is increases to \baseline{one2twoAlgnPsnr} and \baseline{one2sixAlgnPsnr}
respectively.

@@

\section{Experimental Results}\label{sec:results}
We compared $\mu$-path scanning to two scenarios. First, we compare it
to full resolution raster scans at several scan rates in Sec.~\ref{sec:cp_full}.
Second, we compare to course raster scans in Sec.~\ref{sec:cp_subline}.

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/cs_raster_pixel_rows_6-5-2019.svg}
  \caption{Rows of pixels, as indicated by the red lines in
    Fig.~\ref{fig:resultsF1_images}. For clarity, not all images are
    included.}
  \label{fig:pixel_rows}
\end{figure}
For each raster scan and each CS scan, we computed the SSIM and PSNR
 using the 1~Hz raster scan as the master and
also computed the RDI metric. These are plotted against total acquisition time
in Fig.~\ref{fig:time_ssim} and ~\ref{fig:time_psnr}.
Unfortunately, the SSIM and PSNR metrics are not particularly
illuminating. There are several inconsistencies. For example, it makes no sense
the 1~Hz CS scan at 25\% has a \emph{worse} PSNR and SSIM than 5~Hz version;
nor does it make sense that the 8~Hz full raster scan has a worse SSIM and PSNR than
the 10~Hz full raster scan.

@@

\section{Conclusions}\label{sec:conclusions}
The main conclusion of this paper is that with CS, it seems that we
can acquire images faster yet with less specimen damage
than is possible while raster scanning a full image.
On the other hand, it appears that reconstructing a course raster scan makes
more sense for faster scan rates. This is notable, since a course raster scan only
requires an additional step in post-processing. It remains to be seen if this conclusion
holds for a wide range of instruments.
